---
title: "Reproducible Research: Peer Assessment 1"
author: Sarah Weadon
output: 
  html_document:
    keep_md: true
---


### Loading and preprocessing the data

Some text before my code

```{r message = FALSE}
library(readr)
library(dplyr)
library(ggplot2)
```

```{r}
activity.data <- read_csv(unz(list.files()[1], "activity.csv"))
```
Some text after my code


### What is mean total number of steps taken per day?

```{r}
n.days <- n_distinct(activity.data$date)
ttl.steps <- sum(activity.data$steps, na.rm = T)
avg.daily.steps <- ttl.steps / n.days

```

The subject took `r format(ttl.steps, big.mark = ",")` steps over the `r n.days` day period. This is an average of `r format(avg.daily.steps, big.mark = ",")` steps per day.

### What is the average daily activity pattern?

```{r}
daily.ttl <- activity.data %>% 
  group_by(date) %>% 
  summarize(steps = sum(steps, na.rm = T))

ggplot(daily.ttl, aes(steps)) +
  geom_histogram(binwidth = 2500)#, bins = NULL, na.rm = FALSE,
  #show.legend = NA, inherit.aes = TRUE)
```


### Imputing missing values

```{r}

```


### Are there differences in activity patterns between weekdays and weekends?

```{r}

```
