---
title: "Reproducible Research: Peer Assessment 1"
author: Sarah Weadon
output: 
  html_document:
    keep_md: true
---


### Loading and preprocessing the data

Some text before my code

```{r message = FALSE}
library(readr)
library(dplyr)
library(ggplot2)
```

```{r}
activity.data <- read_csv(unz(list.files()[1], "activity.csv"))
```
Some text after my code


### What is mean total number of steps taken per day?

```{r}
total steps per days, histogram, mean, median

n.days <- n_distinct(activity.data$date)
ttl.steps <- sum(activity.data$steps, na.rm = T)
avg.daily.steps <- ttl.steps / n.days

daily.ttl <- activity.data %>% 
  group_by(date) %>% 
  summarize(steps = sum(steps, na.rm = T))

daily.avg <- mean(daily.ttl, na.rm = T) # how does this compare to above?
daily.median <- median(daily.ttl, na.rm = T)

ggplot(daily.ttl, aes(steps)) +
  geom_histogram(binwidth = 2500)

```

The subject took `r format(ttl.steps, big.mark = ",")` steps over the `r n.days` day period. This is an average of `r format(avg.daily.steps, big.mark = ",")` steps per day.

### What is the average daily activity pattern?

```{r}
interval.avg <- activity.data %>%
    group_by(interval) %>%
    summarize(avg.steps = mean(steps, na.rm = T)
    
# compare with above
interval.avg2 <- aggregate(steps ~ interval, activity.data, function(x) mean(x, na.rm = T)

with(interval.avg, plot(interval, avg.steps, type = "l")) # add labels and make pretty
```
The 5 minute interval __ to __ was, on average, the period with the most steps.

### Imputing missing values

```{r}
ttl.missing <- sum(is.na(activity.data$steps))
create a new data set with missing values filled in
repeat analyses from part 1

```
Explain method for filling in missing values
discuss impact of filling in missing values

### Are there differences in activity patterns between weekdays and weekends?

```{r}
create factor: weekday/weekend
repeat part 2 for weekdays and for weekends
```
