---
title: "Reproducible Research: Peer Assessment 1"
author: Sarah Weadon
output: 
  html_document:
    keep_md: true
---

### Introduction

This report looks at step activity collected from midnight October 1, 2012, through 11:59 p.m. November 30, 2012, by an anonomyous individual wearing a personal activity tracking device. Activity was recorded as a number of steps taken in 5-minute intervals.

Four R packages are required for the code included throughout this report. Messages that may appear when loading these packages have been supressed here.

```{r message = FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(stringr)
```

### Loading and preprocessing the data

No preprocessing of the data was required. However, an interval start time and end time in the format of hh:mm were created for graphing convenience. Intervals are 5-minutes each, starting as 0 (midnight), so start and end times follow the pattern illustrated in the table below.

interval | start.time | end.time
---|---|---
0  | 00:00 | 00:04
5  | 00:05 | 00:09
10 | 00:10 | 00:14
15 | 00:15 | 00:19

```{r}
activity.data <- read_csv(unz("activity.zip", "activity.csv")) %>%
  mutate(start.time = str_pad(interval, 4, "left", "0"),
         start.time = paste0(substr(start.time, 1, 2), ":", substr(start.time, 3, 4)),
         end.time = str_pad((interval + 4), 4, "left", "0"),
         end.time = paste0(substr(end.time, 1, 2), ":", substr(end.time, 3, 4)))
```

### What is mean total number of steps taken per day?

```{r}
daily.ttl <- activity.data %>% 
  group_by(date) %>% 
  summarize(steps = sum(steps, na.rm = T))

daily.avg <- mean(daily.ttl$steps)
daily.med <- median(daily.ttl$steps)
```

The subject took `r format(sum(daily.ttl$steps), big.mark = ",")` steps over the course of `r n_distinct(daily.ttl$date)` days. This is an average of `r format(round(daily.avg, 2), big.mark = ",")` steps per day. The median number of steps per day was `r format(daily.med, big.mark = ",")`. The difference between the average and median number of steps per day indicate the data is skewed. Exploring the data further, the histogram below shows `r ` percent of days had 0 to 999 total steps recorded. This indicated a potential problem with the data which is explored and addressed in the "Imputing missing values" section below.

```{r}
ggplot(daily.ttl, aes(x = steps)) +
  geom_histogram(binwidth = 1000, fill = "#4271AE", color = "#1F3552") +
  scale_x_continuous(name = "Total Steps", 
                     breaks = seq(0, 22000, 2000),
                     label = function(x) {format(x, big.mark = ",")}) +
  scale_y_continuous(name = "Number of Days", 
                     breaks = seq(0, 20, 2)) +
  ggtitle("Activity Steps per Day") +
  theme(panel.grid.major.y = element_line(size = 2, color = "gray"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        plot.margin = unit(c(1,2,1,2), "lines"),
        plot.title = element_text(size = 22, face = "bold", 
                                  margin = margin(0,0,20,0)),
        text = element_text(size = 18),
        axis.ticks = element_line(linetype = 0),
        axis.title.x = element_text(margin = margin(20,0,0,0)),
        axis.text.x = element_text(size = 14, margin = margin(-15,0,0,0),
                                   hjust = 1),
        axis.title.y = element_text(margin = margin(0,20,0,0)),
        axis.text.y = element_text(size = 14))
```

### What is the average daily activity pattern?

```{r}
interval.avg <- activity.data %>%
    group_by(interval) %>%
    summarize(avg.steps = mean(steps, na.rm = T))

max.interval <- interval.avg %>% left_join(activity.data, by = "interval") %>%
  select(interval, start.time, end.time, avg.steps) %>% arrange(desc(avg.steps)) %>% slice(1) 

ggplot(interval.avg, aes(interval)) +
  geom_line(aes(y = avg.steps)) +
  scale_x_continuous(name = "Time of Day",
                     breaks = seq(0, 2300, 100),
                     limits = c(0, 2355),
                     label = function(x) { 
                       x <- str_pad(x, 4, "left", "0")
                       x <- paste0(substr(x, 1, 2), ":", substr(x, 3, 4))}) +
  scale_y_continuous(name = "Average Number of Steps", 
                     breaks = seq(0, 250, 50)) +
  ggtitle("ADD TITLE") +
  theme(panel.grid.major.y = element_line(size = 2, color = "gray"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        plot.margin = unit(c(1,2,1,2), "lines"),
        plot.title = element_text(size = 22, face = "bold", 
                                  margin = margin(0,0,20,0)),
        text = element_text(size = 18),
        axis.ticks = element_line(linetype = 0),
        axis.title.x = element_text(margin = margin(20,0,0,0)),
        axis.text.x = element_text(size = 14, margin = margin(-15,0,0,0),
                                   hjust = 1),
        axis.title.y = element_text(margin = margin(0,20,0,0)),
        axis.text.y = element_text(size = 14))
```

The 5 minute interval `r max.interval$start.time` to `r max.interval$end.time` was, on average, the period with the most steps.

### Imputing missing values

```{r}
ttl.missing <- sum(is.na(activity.data$steps))

activity.data <- activity.data %>% mutate(dow = wday(date))
dow.interval.avgs <- activity.data %>% 
    filter(!is.na(steps)) %>% 
    group_by(dow, interval) %>% 
    summarize(avg.steps = mean(steps))

mod.activity.data <- activity.data %>% 
    left_join(dow.interval.avgs, by = c("dow", "interval")) %>%
    mutate(steps = ifelse(is.na(steps), avg.steps, steps) %>%
    select(-avg.steps)
#create a new data set with missing values filled in
#repeat analyses from part 1
mod.daily.ttl <- mod.activity.data %>% 
  group_by(date) %>% 
  summarize(steps = sum(steps, na.rm = T))

mod.daily.avg <- mean(mod.daily.ttl$steps)
mod.daily.med <- median(mod.daily.ttl$steps)

```

There were `r ttl.missing` intervals with missing step data. Intervals with missing data are different from intervals with zero steps recorded.

Missing values were filled in by using avg number of steps in that time period on that day of the week. For example, a missing value for the interval 21:10-21:14 on a Tuesday would be filled in with the average steops for the interval 21:10 - 21:14 on all other Tuesdays without missing values.

# discuss impact of filling in missing values
As you can see from the historam below...

```{r}
ggplot(mod.daily.ttl, aes(x = steps)) +
  geom_histogram(binwidth = 1000, fill = "#4271AE", color = "#1F3552") +
  scale_x_continuous(name = "Total Steps", 
                     breaks = seq(0, 22000, 2000),
                     label = function(x) {format(x, big.mark = ",")}) +
  scale_y_continuous(name = "Number of Days", 
                     breaks = seq(0, 20, 2)) +
  ggtitle("Activity Steps per Day") +
  theme(panel.grid.major.y = element_line(size = 2, color = "gray"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        plot.margin = unit(c(1,2,1,2), "lines"),
        plot.title = element_text(size = 22, face = "bold", 
                                  margin = margin(0,0,20,0)),
        text = element_text(size = 18),
        axis.ticks = element_line(linetype = 0),
        axis.title.x = element_text(margin = margin(20,0,0,0)),
        axis.text.x = element_text(size = 14, margin = margin(-15,0,0,0),
                                   hjust = 1),
        axis.title.y = element_text(margin = margin(0,20,0,0)),
        axis.text.y = element_text(size = 14))
```

### Are there differences in activity patterns between weekdays and weekends?

```{r}
#create factor: weekday/weekend
#repeat part 2 for weekdays and for weekends
mod.activity.data <- mod.activity.data %>%
    mutate(week.weekend = as.factor(ifelse(dow %in% c(1, 7), "weekend", "weekday")))

mod.interval.avg <- mod.activity.data %>%
    group_by(interval) %>%
    summarize(avg.steps = mean(steps, na.rm = T))

mod.max.interval <- mod.interval.avg %>% left_join(mod.activity.data, by = "interval") %>%
  select(interval, start.time, end.time, avg.steps) %>% arrange(desc(avg.steps)) %>% slice(1) 

ggplot(mod.?interval.avg, aes(interval)) +
  geom_line(aes(y = avg.steps)) +
  scale_x_continuous(name = "Time of Day",
                     breaks = seq(0, 2300, 100),
                     limits = c(0, 2355),
                     label = function(x) { 
                       x <- str_pad(x, 4, "left", "0")
                       x <- paste0(substr(x, 1, 2), ":", substr(x, 3, 4))}) +
  scale_y_continuous(name = "Average Number of Steps", 
                     breaks = seq(0, 250, 50)) +
  ggtitle("ADD TITLE") +
  theme(panel.grid.major.y = element_line(size = 2, color = "gray"),
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(), 
        panel.background = element_blank(), 
        plot.margin = unit(c(1,2,1,2), "lines"),
        plot.title = element_text(size = 22, face = "bold", 
                                  margin = margin(0,0,20,0)),
        text = element_text(size = 18),
        axis.ticks = element_line(linetype = 0),
        axis.title.x = element_text(margin = margin(20,0,0,0)),
        axis.text.x = element_text(size = 14, margin = margin(-15,0,0,0),
                                   hjust = 1),
        axis.title.y = element_text(margin = margin(0,20,0,0)),
        axis.text.y = element_text(size = 14))
```
